### 2021-04-20 Code/Exploratory_Analysis.Rmd
commit_EH

---
title: "Exploratory_Analysis"
author: "Elise Harrigan"
date: "4/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up}

getwd()

# Load Packages
library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(leaflet)
library(mapview)
library(raster)
library(sp)
library(rgdal)

#set theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

##load in data - change this to pull from the processed file
tracklogs <- read.csv("../Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("../Data/Processed/IV_wpts_processed.csv")

```
```{r explore the data}
str(tracklogs)
names(tracklogs)
head(tracklogs$lat)
head(tracklogs$lon)

#looking for CRS
st_crs(tracklogs)

```

```{r}
## convert .csv to .shp

 writeOGR(tracklogs, ".", "..Data/Processed/IV_tracklogs_processed.csv", 
           driver = "ESRI Shapefile")

#Read in shapefile into an sf dataframe 
tracklogs_sf <- st_read("../Data/Processed/IV_tracklogs_processed.csv")

#4. Reveal the CRS of the counties features
st_crs(NE_counties_sf)

#5. Plot the data
NE_map <- mapview(NE_counties_sf,
         crs = 4269, 
         grid = FALSE)
NE_map

```