### 2021-04-20 Code/Data_Processing.Rmd
Data Processing updates

---
title: "Data_Processing"
author: "Katie Krejsa"
date: "4/16/2021"
output: pdf_document
---

## Data Setup

```{r}
# set/check working directory
getwd()

# Import Datasets
IV_tracklogs <- read.csv("./Data/Raw/IV_tracklogs.csv")
IV_wpts <- read.csv("./Data/Raw/IV_wpts.csv")

# Load Packages
library(tidyverse)
library(lubridate)
library(plyr)
library(stringr)
library(tidyr)

```

## Data Wrangling

```{r}
# Format date as date class in both datasets
IV_tracklogs$date <- as.Date(IV_tracklogs$date, format = "%Y-%m-%d")
IV_wpts$date <- as.Date(IV_wpts$date, format = "%Y-%m-%d")

# Removing the "z" from the wpts dataset
IV_wpts$time1 <- substr(IV_wpts$time, 1, nchar(IV_wpts$time)-1)

# Finding all of the unique dates in the IV_wpts dataset
wpts_dates <- unique(IV_wpts$date)

# Filtering the IV_tracklogs dataset to that it only includes the unique dates found in the IV_wpts dataset
IV_tracklogs_filtered <- IV_tracklogs[IV_tracklogs$date %in% wpts_dates, ]

# Checking that the unique dates in the IV_tacklogs dataset match the unique dates in the IV_wpts dataset
tracklogs_filtered_dates <- unique(IV_tracklogs_filtered$date)
setequal(wpts_dates, tracklogs_filtered_dates)

# Combining the date and time1 in the IV_wpts dataset
IV_wpts$date_time <- as.POSIXct(paste(IV_wpts$date, IV_wpts$time1), format="%Y-%m-%d %H:%M:%S")

# Combining the date and time in the IV_tracklogs_flitered dataset
IV_tracklogs_filtered$date_time <- as.POSIXct(paste(IV_tracklogs_filtered$date, IV_tracklogs_filtered$time), format="%Y-%m-%d %H:%M:%S")

# arranging the column date_time in IV_wpts dataset by ascending order
IV_wpts_ascending <- arrange(IV_wpts, date_time)

# arranging the column date_time in IV_tracklogs_filtered dataset by ascending order
IV_tracklogs_ascending <- arrange(IV_tracklogs_filtered, date_time)

# Creating a column in IV_tracklogs_ascending called "use"
IV_tracklogs_ascending$use <- NA

# Assigning cmt column values to the IV_tracklogs_ascending dataset
for (i in 1:nrow(IV_wpts_ascending)-1){
  IV_tracklogs_ascending$use[IV_tracklogs_ascending$date_time >= IV_wpts_ascending$date_time[i] & IV_tracklogs_ascending$date_time <= IV_wpts_ascending$date_time[i+1]] <- IV_wpts_ascending$cmt[i]
  }

# filtering out all of the rows in the IV_tracklogs_ascending dataset that have "NA" and blank in the "use" column so we are left with only entries that have values in the "use" column
IV_tracklogs_ascending <-
  IV_tracklogs_ascending %>% 
  filter(!is.na(use))

# If the "use" column in IV_tracklogs_ascending is blank then the details previously recorded remained the same; Therefore, need to fill in blanks with previously recorded value; Need to fill cells of empty strings with NA first before filling
IV_tracklogs_ascending <-
  IV_tracklogs_ascending %>% 
  mutate(use = if_else(use=="", NA_character_,use)) %>% 
  fill(use, .direction = "down")

# Separating the characters and numbers in the "use" column of the IV_tracklogs_ascending dataset to make two new columns
IV_tracklogs_ascending$use_cha <- (str_extract(IV_tracklogs_ascending$use, "[aA-zZ]+"))
IV_tracklogs_ascending$use_num <- as.numeric(str_extract(IV_tracklogs_ascending$use, "[0-9]+"))

# Adding the "name" column from the IV_wpts_ascending dataset to the IV_tracklogs_ascending dataset
IV_tracklogs_ascending$name <- NA

for (i in 1:nrow(IV_wpts_ascending)-1){
  IV_tracklogs_ascending$name[IV_tracklogs_ascending$date_time >= IV_wpts_ascending$date_time[i] & IV_tracklogs_ascending$date_time <= IV_wpts_ascending$date_time[i+1]] <- IV_wpts_ascending$name[i]
  }

```

## Summarizing Data Structure

```{r}
# Ranges of dates
range(IV_tracklogs$date)
range(IV_wpts$date)

# Range and central tendancies of trail widths
range(IV_tracklogs_ascending$use_num)
mean(IV_tracklogs_ascending$use_num)
median(IV_tracklogs_ascending$use_num)

# Range of lattitude and longitude
range(IV_tracklogs$lat)
mean(IV_tracklogs$lat)
median(IV_tracklogs$lat)

range(IV_tracklogs$lon)
mean(IV_tracklogs$lon)
median(IV_tracklogs$lon)

# unique values for use_cha
unique(IV_tracklogs_ascending$use_cha)

```

## Saving Processed Datasets

```{r}

write.csv(IV_tracklogs_ascending, row.names = FALSE, file = "./Data/Processed/IV_tracklogs_processed.csv")

write.csv(IV_wpts_ascending, row.names = FALSE, file = "./Data/Processed/IV_wpts_processed.csv")

```