### 2021-04-21 Code/Exploratory_Analysis.Rmd
Exploratory_analysis1

---
title: "Exploratory_Analysis"
author: "Elise Harrigan"
date: "4/20/2021"
output: html_document
geometry: margin=2.54cm
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up}

getwd()

# Load Packages
library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(leaflet)
library(mapview)
library(raster)
library(sp)
library(rgdal)
library(cowplot)
#install.packages("rnaturalearth")
#install.packages("rnaturalearthdata")
library(rnaturalearth)

#set theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom", legend.justification = "right")
theme_set(mytheme)

##load in data - change this to pull from the processed file
tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")

```

```{r explore the data}
str(tracklogs)
names(tracklogs)
head(tracklogs$lat)
head(tracklogs$lon)
summary(tracklogs)

#looking for CRS
st_crs(tracklogs)
#none found but Amelia said WGS 84. Googled UTM projection and is in Zone 32S

tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")

```

```{r}
#subset based on quality of trail
Poor_trails <- tracklogs %>%
  filter(use_cha == "P")

Med_trail <- tracklogs %>% 
  filter(use_cha == "M")

##W not H??
Heavy_trail <- tracklogs %>%
  filter(use_cha == "W")

```

```{r how many P, M, W trails}

P_num <- ggplot(tracklogs, aes(y = use_cha)) +
  geom_bar() + mytheme

print(P_num)

```
__Condition of Trail __      | __Total Amount__
--------------------         |---------------------------
Poor (P)                     | 27,218
Medium (M)                   | 13,666
Heavy (H)                    | 26

```{r}
##plot in ggplots with lat, long

#world is from the rnaturalearth package. gives background for country
world <- ne_countries(scale = "medium", returnclass = "sf")

##ggplot using world map under
Trail_gg <- ggplot(data = world) +
     geom_sf() +
    geom_point(data = tracklogs, aes(x = lon, y = lat, color = use_cha), size = .4)  + 
    coord_sf(xlim=c(12.20, 12.90), ylim=c(-0.5, .8), expand = TRUE) +
  ggtitle("Elephant Trail Points") + 
  xlab("Longitude") + ylab("Latitude") + 
  scale_x_continuous(breaks=seq(12, 13, by = .5)) +
 mytheme
print(Trail_gg)

## sarah's theme code
##theme(panel.background = element_rect(fill = "white", colour = "black"))

##PLOT IN MAPVIEW for one date

##filter for dates from 02-28-2018 -- 03-10-2018
subset_trails <- tracklogs %>%
  filter(date == "2018-02-28")

#CRS=WGS84 = 4326 or UTM zone 32S = 32732

#convert to sf and include crs (I used UTM zone 32S and coord columns)
subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lat", "lon"))

#transform to UTM
trails_UTM <- st_transform(subset_sf, crs=32633)

#Plot the data to see the trails - the projection is off by a little 
Trail_map1 <- mapview(trails_UTM)
  Trail_map1

```

## ANALYSIS
```{r plots by width}

name_plots <- ggplot(tracklogs, aes(x = use_cha, y = use_num)) +
  geom_boxplot(fill = "darkblue") + ggtitle("Trail type by width") + xlab("Type of Trail") + ylab("Width of trail") + mytheme

print(name_plots)

```

```{r boxplots trail by widths}

##boxplot width of poor trails
PT_plot <- ggplot(Poor_trails, aes(x=use_cha, y=use_num)) +
  geom_boxplot(fill = "seagreen") + 
  ggtitle("Poor Troddened Trail by widths") + 
  xlab("Amount of use") + 
  ylab("Width of trail") + ylim(35, 55) + mytheme
print(PT_plot)

##boxplot width of medium trails
MT_plot <- ggplot(Med_trail, aes(x=use_cha, y = use_num)) + 
  geom_boxplot(fill = "seagreen3") +
  ggtitle("Medium Troddened Trail by widths") + 
  xlab("Amount of use") + 
  ylab("Width of trail") + mytheme

print(MT_plot)

##boxplot width of heavy trails
HT_plot <- ggplot(Heavy_trail, aes(x=use_cha, y = use_num)) + 
  geom_boxplot(fill = "seagreen4") +
  ggtitle("Heavy Troddened Trail by widths") + 
  xlab("Amount of use") + 
  ylab("Width of trail") + mytheme

print(HT_plot)

##stack the boxplots
tstack<- plot_grid(
  PT_plot,
  MT_plot,
  HT_plot,
align = 'vh', 
hjust = -1, 
ncol = 1)

tstack

##boxplot frequency - number of days a trail was used

ggplot(tracklogs, aes(x=date, fill = use_cha)) + 
  geom_bar() + 
  scale_x_date(date_breaks ="10 days", date_labels = "%Y-%m-%d", 
          guide = guide_axis(angle = 45)) + 
  ggtitle("Amount of Trails by day") +
  mytheme

```

```{r model - Analysis}
#is the width of the trail significant on the frequency of use?

#do any of the variables (lat, lon, use_num) explain use_cha
mod_1 <- lm(data = tracklogs, factor(use_cha) ~ lat + lon + use_num)
  summary(mod1)

```